### 2013-07-06 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
adding msod scaling section

\subsection{Multiple Set Domain Decomposition}
\label{subsec:msod_decomposition}

Given that overlap was not successful in boosting parallel
efficiencies, we next look to apply replication to the problem in the
form of multiple sets. Before performing parallel scaling studies
using multiple sets, several aspects of the replicated parallel
problem must be considered in order to effectively craft the studies
and interpret their results. First is the question of how many
stochastic histories to use to generate the correction vector in the
MCSA iteration. For this work we consider two choices; the split case
and the fully replicated case. As an example these cases, consider a
single set problem where 20,000 histories are used in the Monte Carlo
solve. If we choose the split case, then solving the same problem with
2 sets would give 10,000 histories in each set and 4 sets would give
5,000 histories in each set. Splitting the histories in this way
provides the same global number of histories used to compute the
correction in the single set problem. We expect to reduce the time to
solution by doing this as each individual set solve will occur faster
in an almost linear fashion with the reduced number of histories. For
the replicated history case, each additional set in the problem will
use just as many histories as the single set case. Using 2 sets with
fully replicated histories for our example problem then has each set
computing 20,000 histories with 40,000 global histories in the
computation. For the 4 set case, the number of global histories is
increased to 80,000. We also expect this type of replication to reduce
the time to solution as the extra global histories should reduce the
stochastic uncertainty in the correction vector and reduce the overall
number of MCSA iterations required to converge.

@@

@@

Based on the analysis of the pure domain decomposed MCSA algorithm and
subsequently the MSOD implementation, one may ask then ask if there is
any need to communicate histories from domain-to-domain at all. Overlap
was not successful because so few histories needed to be communicated
in the first place that keeping them on-process resulted in more work
than the communications savings that occurred. If that is true, then
perhaps without any overlap or communication, just doing local Monte
Carlo may be satisfactory for accelerating the MCSA iteration. In this
case, histories near the boundaries of the domain may be truncated
before reaching the weight cutoff if the reach the domain boundary
first. This truncation has the potential to increase the stochastic
uncertainty of the Neumann-Ulam result and therefore increase the
number of MCSA iterations required to converge. In addition, the
scalability of the method will then be purely limited by the parallel
matrix and vector operations required to implement the remainder of
the algorithm.

@@

Of additional concern for the subdomain method is the potential
decrease in iterative performance due to the fact that histories near
the boundary will potentially have their random walk truncated, thus
decreasing the information available to compute the correction
vector. For both the weak and strong scaling study, all calculations
converged in 22 or in only a few cases 21 iterations for both the
subdomain solve and the solve with communication giving an algorithmic
efficiency of 1.048 for all cases. This small improvement in iteration
count was not observed to be a function of problem size, number of
processors, or whether the subdomain method was used. Therefore, there
is no reason to expect overlap to improve the iterative or scaling
performance of the method as iterative performance is maintained
without it. Overlap is not effective because histories relinquish most
of their useful information within the first few steps (which is why
we can use such a large weight cutoff and there is an insensitivity to
weight cutoff as shown in
Figure~\ref{fig:estimator_wc_iters}). Keeping them around longer with
overlap to do extra work in the local domain will not improve
iterative performance and thus increasing compute times.

>>>>

\subsection{Multiple Set Domain Decomposition}
\label{subsec:ms_decomposition}

We next look to apply replication to the problem in the form of
multiple sets. Before performing parallel scaling studies using
multiple sets, several aspects of the replicated parallel problem must
be considered in order to effectively craft the studies and interpret
their results. First is the question of how many stochastic histories
to use to generate the correction vector in the MCSA iteration. For
this work we consider two choices; the split case and the fully
replicated case. As an example these cases, consider a single set
problem where 20,000 histories are used in the Monte Carlo solve. If
we choose the split case, then solving the same problem with 2 sets
would give 10,000 histories in each set and 4 sets would give 5,000
histories in each set. Splitting the histories in this way provides
the same global number of histories used to compute the correction in
the single set problem. We expect to reduce the time to solution by
doing this as each individual set solve will occur faster in an almost
linear fashion with the reduced number of histories. For the
replicated history case, each additional set in the problem will use
just as many histories as the single set case. Using 2 sets with fully
replicated histories for our example problem then has each set
computing 20,000 histories with 40,000 global histories in the
computation. For the 4 set case, the number of global histories is
increased to 80,000. We also expect this type of replication to reduce
the time to solution as the extra global histories should reduce the
stochastic uncertainty in the correction vector and reduce the overall
number of MCSA iterations required to converge.

@@

\subsection{Multiple-Set Overlapping-Domain Decomposition}
\label{subsec:msod_decomposition}

@@

Based on the analysis of the pure domain decomposed MCSA algorithm and
subsequently the MSOD implementation, one may ask then ask if there is
any need to communicate histories from domain-to-domain at
all. Overlap was not as successful because so few histories needed to
be communicated in the first place that keeping them on-process
resulted in more work than the communications savings that
occurred. If that is true, then perhaps without any overlap or
communication, just doing local Monte Carlo may be satisfactory for
accelerating the MCSA iteration. In this case, histories near the
boundaries of the domain may be truncated before reaching the weight
cutoff if the reach the domain boundary first. This truncation has the
potential to increase the stochastic uncertainty of the Neumann-Ulam
result and therefore increase the number of MCSA iterations required
to converge. In addition, the scalability of the method will then be
purely limited by the parallel matrix and vector operations required
to implement the remainder of the algorithm.

@@

Of additional concern for the subdomain method is the potential
decrease in iterative performance due to the fact that histories near
the boundary will potentially have their random walk truncated, thus
decreasing the information available to compute the correction
vector. For both the weak and strong scaling study, all calculations
converged in 22 or in only a few cases 21 iterations for both the
subdomain solve and the solve with communication giving an algorithmic
efficiency of 1.048 for all cases. This small improvement in iteration
count was not observed to be a function of problem size, number of
processors, or whether the subdomain method was used. Therefore, there
is no reason to expect overlap to improve the iterative or scaling
performance of the method as iterative performance is maintained
without it. Overlap is not as effective because histories relinquish
most of their useful information within the first few steps (which is
why we can use such a large weight cutoff and there is an
insensitivity to weight cutoff as shown in
Figure~\ref{fig:estimator_wc_iters}). Keeping them around longer with
overlap to do extra work in the local domain will not improve
iterative performance and thus increasing compute times.
